---
title: "Plots of predicted dropout probability"
output: pdf_document
---

```{r echo=FALSE, message = FALSE}
library(ggplot2)
library(umx)
library(dplyr)
library(here)
source(here("code/table_helper_functions.R"))
```

```{r echo=FALSE, message = FALSE}
models <- c("AO-2S-PA", "MIM-2S-PA", "AO", "MIM", "Mean Score")

df <- readRDS(here("rds/score_df_continuous.rds"))
df$dropout_f <- mxFactor(df$dropout, levels = c(0, 1))
df$ELS <- 0
df$ELS[df$sample=="ELS"] <- 1

# prepare dataframes with the standardized self_efficacy variables 
df_ao <- cbind(df[, c("dropout_f", "ELS")], "self_efficacy" = df$self_efficacy_ao_s)
df_mim <- cbind(df[, c("dropout_f", "ELS")], "self_efficacy" = df$self_efficacy_mim_s)
df_mn <- cbind(df[, c("dropout_f", "ELS")], "self_efficacy" = df$self_efficacy_mn_s)
```

```{r results= 'hide', warning = FALSE, message = FALSE, echo = FALSE}
# 1 - NO RELIABILITY ADJUSTMENT (PA) ####
umx_ao <- umxRAM("
    dropout_f ~ self_efficacy + ELS
    dropout_f ~ 0 * 1
    dropout_f ~~ 1 * dropout_f", 
    name = "probit regression (AO)", data = df_ao, tryHard = "yes")
umx_mim <- umxRAM(
   "dropout_f ~ self_efficacy + ELS
    dropout_f ~ 0 * 1
    dropout_f ~~ 1 * dropout_f", 
    name = "probit regression (MIM)", data = df_mim, tryHard = "yes")
umx_mn <- umxRAM(
   "dropout_f ~ self_efficacy + ELS
    dropout_f ~ 0 * 1
    dropout_f ~~ 1 * dropout_f", 
    name = "probit regression (Mean Scores)", data = df_mn, tryHard = "yes")

# 2 - WITH RELIABILITY ADJUSTMENT (2S-PA) ####
# drop rows with missing reliability
df_a <- cbind(df[!is.na(df$approx_rel), ], 
              "self_efficacy" = df[!is.na(df$approx_rel),"self_efficacy_ao_s"],
              "self_efficacy_ev" = df[!is.na(df$approx_rel),"approx_ev"])
df_p <- cbind(df[!is.na(df$partial_rel), ], 
              "self_efficacy" = df[!is.na(df$partial_rel),"self_efficacy_mim_s"], 
              "self_efficacy_ev" = df[!is.na(df$partial_rel),"partial_ev"])

umx_tspa_ao <- umxRAM("2spa (AO)",
  umxPath(c("math", "ELS"), to = "dropout_f"), # main effects
  # loading = reliability (as definition variable) = 1 for Bartlett FS
  umxPath("math", to = "self_efficacy", values = 1, free = FALSE),
  # error variance (as definition variable), fs_SE^2
  umxPath(var = "self_efficacy", labels = "data.self_efficacy_ev", free = FALSE),
  umxPath(unique.pairs = c("math", "ELS")), # covariances of predictors
  umxPath(means = c("math", "ELS")),  # means of predictors
  umxPath(v1m0 = "dropout_f"), # for model identification, dropout_f ~ N(0, 1)
  data = df_a, tryHard = "yes"
)
umx_tspa_mim <- umxRAM("2spa (MIM)",
  umxPath(c("math", "ELS"), to = "dropout_f"), # main effects
  # loading = reliability (as definition variable) = 1 for Bartlett FS
  umxPath("math", to = "self_efficacy", values = 1, free = FALSE),
  # error variance (as definition variable)
  umxPath(var = "self_efficacy", labels = "data.self_efficacy_ev", free = FALSE),
  umxPath(unique.pairs = c("math", "ELS")), # covariances of predictors
  umxPath(means = c("math", "ELS")), # means of predictors
  umxPath(v1m0 = "dropout_f"), # for model identification, dropout_f ~ N(0, 1)
  data = df_p, tryHard = "yes"
)
```


```{r echo=FALSE}
cols <- c("#F26178", "#FFCC00")

# Probit regression predicted probability plot by sample
plot_dropout <- function(umx_out, approach, seq_f, seq_t, seq_by) {

  if(approach == "AO-2S-PA" | approach == "MIM-2S-PA") {
    terms <- c("dropout_f_dev1", "math_to_dropout_f", "ELS_to_dropout_f")
  } else {
    terms <- c("dropout_f_dev1", "self_efficacy_to_dropout_f", "ELS_to_dropout_f")
  }
  
  # get path coefficients 
  coefs <- get_path_ests_umx(terms, umx_out)
  
  # make predictions for self-efficacy scores of seq(-2, 2, 0.1)
  new_self_eff <- seq(seq_f, seq_t, seq_by)
  # construct design matrices w columns fosr intercept, self-efficacy, ELS
  new_mat_els <- cbind(1, new_self_eff, rep(1, length(new_self_eff)))
  new_mat_hsls <- cbind(1, new_self_eff, rep(0, length(new_self_eff)))
  
  # predictions: Xnew %*% beta_hat
  els_preds <- new_mat_els %*% coefs[,1]
  hsls_preds <- new_mat_hsls %*% coefs[,1]
  
  # Var(predicted y) = xnew %*% V(beta-hat) %*% t(xnew)
  # SE = sqrt(Var(predicted y))
  v_betahat <- vcov(umx_out)[terms, terms]
  v_betahat[2:3, 1] <- -v_betahat[2:3, 1] # flip the signs 
  v_betahat[1, 2:3] <- -v_betahat[1, 2:3]
  
  se_preds_els <- apply(new_mat_els, 1, function(row) {
    row <- matrix(row, nrow = 1)
    sqrt(row %*% v_betahat %*% t(row))
  })
  se_preds_hsls <- apply(new_mat_hsls, 1, function(row) {
    row <- matrix(row, nrow = 1)
    sqrt(row %*% v_betahat %*% t(row))
  })
  
  lb_els <- els_preds - 1.96 * se_preds_els
  ub_els <- els_preds + 1.96 * se_preds_els
  lb_hsls <- hsls_preds - 1.96 * se_preds_hsls
  ub_hsls <- hsls_preds + 1.96 * se_preds_hsls
  
  plot(new_self_eff, pnorm(els_preds), type = "n", xlab = "math self-efficacy", 
       ylab = "P(dropout)", ylim = c(0, .2), bty = "l", font.main = 1,
       main = paste0("Predicted Dropout Probability vs. Math Self-Efficacy (", 
                     approach, ")"))
  
  lines(new_self_eff, pnorm(els_preds), type = "l", col = cols[1], lwd = 2)
  lines(new_self_eff, pnorm(lb_els), type = "l", col = cols[1], lwd = 1, lty = 2)
  lines(new_self_eff, pnorm(ub_els), type = "l", col = cols[1], lwd = 1, lty = 2)
  
  lines(new_self_eff, pnorm(hsls_preds), type = "l", col = cols[2], lwd = 2)
  lines(new_self_eff, pnorm(lb_hsls), type = "l", col = cols[2], lwd = 1, lty = 2)
  lines(new_self_eff, pnorm(ub_hsls), type = "l", col = cols[2], lwd = 1, lty = 2)
}
```



```{r echo=FALSE, results='hide', message = FALSE}
plot_dropout(umx_tspa_ao, "AO-2S-PA", -2, 2, 0.1)
```

Note: A probit regression model was used to predict the probability of dropout from sample membership and Bartlett factor scores (FS) for math self-efficacy. Following a two-stage path analysis (2S-PA) approach, FS were computed using the approximate invariance model determined via alignment optimization (AO) in the first stage to correct for measurement noninvariance. The scores were further corrected for unreliability in the second stage. The dotted lines indicate 95% confidence intervals. 

\pagebreak
```{r echo=FALSE, results='hide', message = FALSE}
plot_dropout(umx_tspa_mim, "MIM-2S-PA", -2, 2, 0.1) 
```

Note: A probit regression model was used to predict the probability of dropout from sample membership and Bartlett factor scores (FS) for math self-efficacy. Following a two-stage path analysis (2S-PA) approach, FS were computed using the partial invariance model determined via measuement invariance modeling (MIM) in the first stage to correct for measurement noninvariance. The scores were further corrected for unreliability in the second stage. The dotted lines indicate 95% confidence intervals.

\pagebreak

```{r echo=FALSE, results='hide', message = FALSE}
plot_dropout(umx_ao, "AO", -2, 2, 0.1)
```

Note: A probit regression model was used to predict the probability of dropout from sample membership and Bartlett factor scores (FS) for math self-efficacy. FS were computed using the approximate invariance model determined via alignment optimization (AO). The scores were not corrected for unreliability. The dotted lines indicate 95% confidence intervals.
\pagebreak

```{r echo=FALSE, results='hide', message = FALSE}
plot_dropout(umx_mim, "MIM", -2, 2, 0.1)
```


Note: A probit regression model was used to predict the probability of dropout from sample membership and Bartlett factor scores (FS) for math self-efficacy. FS were computed using the partial invariance model determined via measuement invariance modeling (MIM). The scores were not corrected for unreliability. The dotted lines indicate 95% confidence intervals.
\pagebreak

```{r echo=FALSE, results='hide', message = FALSE}
plot_dropout(umx_mn, "Mean Score", -2, 2, 0.1)  
```
Note: A probit regression model was used to predict the probability of dropout from sample membership and composite mean scores of the math self-efficacy test items. The scores were not corrected for measurement error or unreliability. The dotted lines indicate 95% confidence intervals.
